Compositional equivalences based on Open pNets

### 2020-04-24 2016-Weak-FH-Bisim/WeakBisim.tex
Big changes in use-case texts
Partout sauf les figures !!

\subsubsection*{Substitutions}

@@

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, V, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of states.
\item[$\bullet$]
$s_0 \in S$ is the initial state.
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
\item[$\bullet$] $V$ is a set of global variables for the pLTS,
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
where $\alpha \in\AlgA$ is a parameterised action, $e_b \in
\AlgB$ is a guard, and the variables $x_j$ 
are assigned the expressions $e_j\in \AlgE$.
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then 
% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! V$, 
		$\vars(e_b)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
		$\forall j\!\in\! J .\,\left(\vars(e_j)\!\subseteq\! V\cup\vars(\alpha)\land 
		x_j\!\in V \right)$. %,  and $s= s'\Rightarrow J=\emptyset$.

@@

We now define the structure of pNets, the following definition relies on the definition 
of holes, leaves and sorts formalised below in Definition~\ref{def-sortholeleave}. Informally, holes are process parameters, leaves provide the set of pLTSs at the leaves of the hierarchical structure of a pNet, and sorts give the signature of a pNete, i.e. the actions it exposes.

@@

Fig. \ref{SimpleProt:Impl} shows the pNet model of the protocol implementation. When the $\symb{Medium}$ detects an error (modelled by a local $\tau$ action), it sends back a $\texttt{m\_error}$ message to the $\symb{Sender}$. The $\symb{Sender}$ increments its local error counter $ec$, and resends the message (including $ec$) to the $\symb{Medium}$, that will, eventually, transmit $m,ec$ to the $\symb{Receiver}$. Strictly speaking, as variables are local to each  state, we have two distinct $ec$ variables in the \textit{Sender} pLTS, one in state \texttt{s1}, another one in state \texttt{s2}, the value of the variable is updated along the transitions entering each of the two states.

@@

\subsection{Open Automata}
 Open automata (OA) are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on a set of variables (potentially with only symbolic values). 
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} (OT) over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$, a set $V$ of variables, and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq J$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action of this open transition.
%        We call \emph{source variables} of an OT the union of all variables
%        in the different terms
%        $\beta_j$ and $\alpha$, and the state variables of $s$.
        \Pred\ is a predicate over the  variables $V$ and all variables
        in the different terms
       $\beta_j$ and $\alpha$. \Post\ is a set of 
	assignments that are effective \emph{after the open transition}, they are
        represented as a substitution of the form $({x_k\gets e_k})^{k\in K}$ 
	where $\forall k.\, x_k\in V$, and $e_k$ are expressions
        over the variables $V$ and all variables
        in the different terms
       $\beta_j$ and $\alpha$. Because the variables in $V$ can be in both side of the assignments, the assignments are always applied in a parallel manner. Open transitions are identified
        modulo logical equivalence on their predicate. 
\end{definition}

@@

We take in this article a semantics and logical understanding of these
automata. Open automata are closed by a simple form of refinement that
allows us to refine the predicate, or substitute any free variable by
an expression. More formally, let $\Post$ be any substitution and
$\Pred$ any predicate; suppose $V\cap \dom(\Post)=\emptyset$. Then we have the following
implication:

@@

%\LUDO{I added this paragraph}
Because of the semantic interpretation of open automata, the set of open transition of an open automaton is infinite (for example because every free variable can be renamed). However an open automaton is characterized by a  subset of these open transition which is sufficient to generate, by substitution the other ones. In the following, we will abusively write that we define an ``open automaton'' when we provide only the set of open transitions that is sufficient to generate a proper open automaton by saturating each open transition by all possible substitutions.

@@

We define first states of open pNets as tuples of states. We denote them
 as $\triangleleft\ldots\triangleright$ for distinguishing tuple 
states from other tuples.
\begin{definition}[States of open pNets]\label{def-states}
  A state of an open pNet is a tuple (not necessarily finite) of the
  states of its leaves.

@@

Somehow, this predicate entails a verification of satisfiability in the sense that if the 
predicate $\Predsv$ is not satisfiable, then the transition associated with the 
synchronisation will not occur in the considered state, or will occur with a \False\ precondition which is equivalent.
If the action families do not match or if there is no valuation of
variables such that the above formula can be ensured the predicate is undefined.

@@

 \smallskip\noindent
 \[  OT_1  = \openrule{
      \{\texttt{P}\mapsto \texttt{p-send(m)}\},  [\texttt{m=m'}],
        \{\texttt{s1\_msg}\gets \texttt{m}\}
                      }
    {\ostate{s_0,m_0,r_0} \OTarrow{\nounderline{\texttt{in(m')}}} \ostate{s_1,m_0,r_0}}
    \]

    \smallskip
    The global states here are triples, build as the product of states of the 3 pLTS (remember the holes have no state). The assignment
    $Post$ uses the variable $\texttt{m}$ from the action of hole $\texttt{P}$ to set the
    value of the sender state variable named $\texttt{s1\_msg}$.

@@

 %% \smallskip\noindent
 %% $  OT_2  = \openrule{
 %%   \emptyset; Pred=[send(SV2\_m)=send(s_1\_m);in(SV2\_m)=in(?m)];  \\ Post=\{m_1\_m:=?m\}
 %%                      }
 %%    {\ostate{s_1,m_0,r_0} \OTarrow{\nounderline{send(SV2\_m)}} \ostate{s_2,m_1,r_0}}
 %%    $

@@

 %%     $  OT_2  = \openrule{
 %%      \emptyset; Pred=\True;  \\ Post=\{m_1\_m:=s_1\_m\}
 %%                      }
 %%    {\ostate{s_1,m_0,r_0} \OTarrow{\nounderline{send(s_1\_m)}} \ostate{s_2,m_1,r_0}}
 %%    $

 %%    \smallskip
 %%    Note that the global states shows how the sender and the
 %%    first medium have evolved during this transition, the sender
 %%    moving from $s_1$ to $s_2$, and the first (forward) medium from
 %%    $m_0$ to $m_1$.
 %%    The predicate $Pred$ involves the state variables $s_1\_m$ and
 %%    $s_1\_sb$ of the sender, while the assignment $Post$ sets the
 %%    state variables of the medium.

@@

%% \LUDO{Je sais pas de quoi parle le prochain todo :-( l'exemple?}
%% \TODO{Ceci n'est pas nouveau, n'est-ce pas, est-ce utile pour le lecteur ici ? Ou est-ce qu'on bouge tous ces ``details techniques'' dans une annexe ? Si oui, je refais avec un morceau de ABP}
%% \begin{example} \emph{Using the operational rules to compute
%%     open-transitions}
%%   In Fig. \ref{usingrules:OT2} we show the deduction tree used to construct and prove the 
%%   open transition $OT_2$ of \texttt{EnableCompL} (see example page \pageref{OT:ABP-composed}).
%%   The rule uses \TrUn\ for the $\delta$ transition of $C_3$, for the $l$ transition of $C_4$, then combines the result using the $a_4$ vector of the bottom pNet node, and the $\nounderline{\delta(x)}$ vector of the top node.

%% \begin{figure}[h]
%% \begin{mathpar}
%%   \small
%%   \inferrule
%%     {\inferrule
%%         {0 \xrightarrow {\delta}_{C_3} 1}
%%         {C_3
%%           \models
%%             \openrule{
%%               0 \xrightarrow {\delta}_{C_3} 1,\,
%%               \{\xrightarrow{\delta(x_1)}_P\},\,
%%               v_1=\delta(x_1)}
%%                       {\ostate{0}\OTarrow{v_1}\ostate{1}}
%%         }\\
%% %      \sm{fresh}v         \\
%%       \inferrule%*[right={L_1}]
%%         {
%% %          \sm{fresh}{a_Q} \\
%%           \inferrule
%%               {0 \xrightarrow{l}_{C_4} 0}
%%               % {\ostate{0}\xrightarrow{l}\ostate{0}}
%%               {C_4
%%                 \models
%%                 \openrule{
%%                       0 \xrightarrow l_{C_4} 0,\,
%%                       \sm{Pred}_{C_4}}
%%                       {\ostate{0}\OTarrow{l}\ostate{0}}
%%               }
%%         }
%%         {
%%           \textrm{Q>>R}\models
%%               \openrule
%%                   { 0 \xrightarrow {l}_{C_4} 0,\,
%%                     \{\xrightarrow{acc(x_2)}_Q\},\,
%%                     \ v_2=acc(x_2)
%%                   }
%%                   {\ostate{0}\OTarrow{v_2}\ostate{0}}
%%         }   
%%     }
%%     {
%%      \textrm{P>>(Q>>R)}
%%      \models
%%      \openrule
%%          { 0 \xrightarrow{\delta}_{C_3} 1, \\ 0 \xrightarrow{l}_{C_4} 0,\\
%%            \{\xrightarrow{\delta(x)}_P,\,\xrightarrow{acc(x)}_Q\}, \\
%%             a_3=v_1 \wedge v=a_3 \wedge x_1=x_2
%%            }
%%          {\ostate{00} \OTarrow{v} \ostate{10}}
%%       }\vspace{-4ex}
%% \end{mathpar}
%%   \caption{Proof of transition $OT_2$ (with interaction of processes $P$ and $Q$) for 
%%   ``P>>(Q>>R)'' \TODO{Change for the new example or remove?}}
%%   \label{usingrules:OT2}
%% \end{figure}

%% \end{example}

@@

\subsection*{Example}  Figure~\ref{SimpleProtCounter:SpecOA} shows the open automaton computed from the Specification pNet given in Figure \ref{SimpleProt:Spec}. 
For later references, we give to the transitions of this (strong)
Specification automaton as $SS_i$ while transitions of the
Implementation automaton will be labelled $SI_i$. In the figure we
have annotated each global state with the set of corresponding state variables.

@@

\end{definition}
Rule $\WTUn$ states that it is always possible to do a null $\tau$ transition, where the state is unchanged and the holes perform no action. Rule~$\WTDeux$ states that each open transition can be considered as a weak open transition. The last rule is the most interesting:  Rule~$\WTTrois$ allows any number of $\tau$ transitions before or after a weak open transition. This rules carefully composes predicates, effects, and actions of the holes, indeed in the rule, predicate $\Pred_2$ manipulates variables of $s_1$ that result from the first weak open transition. Their value thus depend of the initial state but also of the effect (as a substitution $\Post_1$) of the first weak open transition. In the same manner, $\Pred_3$ must be applied the joint substitution $\Post_2\bigotimes\Post_1$. Similarly, effects on variables must be applied to obtain the global effect of the composed weak open transition, it must also be applied to observable actions of the holes, and to the global action of the weak open transition.

@@

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.  For reasons of readability, we abbreviate the weak open transitions encoded by $\openrule   {\{\}, True,	() } {s \OTWeakarrow \tau s'}$  as $W_\tau$. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds an $WT_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {\text{b1}\OTWeakarrow \tau \text{b1}}$; $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$;  $\openrule
         {\{\}, True,
			() }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+2) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
        {\text{b1} \OTWeakarrow \tau \text{b1}} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{\nounderline{out(b1\_m,b1\_ec)}}  as $\alpha$, we get the resulting weak open transition between b1 and b0 (as shown in Figure \ref{WOT2}).

@@

 For readability, we did not include the full details of the Implementation weak open transitions in Figure \ref{SimpleProtCounter:ImplWOA2}, they will appear below.
First, we point out that the weak OT loops ($WI_1$,$WI_2$ and $W_\tau$) on state ${000}$ are also present in all other states, we did not repeat them. Then many WOTs are similar, and numbered accordingly as 3, 3a, 3b, 3c and 8, 8a, 8b, 8c respectively: they only differ by the name of the state variables in their respective source or target states.
\end{example}

@@

Now let us give some details about the construction of the weak automaton of the implementation pNet, obtained by application of the weak rules as explained above. We concentrate on weak open transitions $WI_3$ and $WI_4$. Let us denote as $post_n$ the effect (= substitution) of the strong open transitions $OT_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

$post_3 = (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets 0)$

$post_4 = (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}, \texttt{s2\_ec}\gets \texttt{s1\_ec})$

@@

$post_6 = (\texttt{s1\_ec}\gets \texttt{s2\_ec+1})$

@@

$post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (\texttt{s1\_ec}\gets \texttt{s1\_ec}+1)$.
\medskip

So if we denote ${post_{456*}}$ any iteration of this loop, we get ${post_{456*}} = (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n)$ for any $n\ge 0$, and the \Post~ of the weak OT  $WI_{3}$ is:\\
 ${post_{456*}}\shortotimes\ post_3 = (\texttt{s1\_msg}\gets m, \texttt{s1\_ec}\gets n), \forall n\ge 0$ and \Post~ of  $WI_{3a}$  is:\\ $post_4\shortotimes {post_{456*}}\shortotimes\ post_3 = (\texttt{m1\_msg}\gets m, \texttt{m1\_ec}\gets n), \forall n\ge 0$.
\medskip

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\nounderline{\texttt{in(m)}}} 100}
        $$\forall n\ge 0$

$ WI_4(n) = \openrule
         {\{\}, True, 
   (\texttt{m1\_msg}\!\gets\! \texttt{s1\_msg}, \texttt{m1\_ec}\!\gets\! \texttt{s1\_ec}+n, \texttt{s2\_ec}\!\gets\! \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
        $$\forall n\ge 0$

$ WI_{456}(n) = \openrule
         {\{\}, True, 
    (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n)}
  {100 \OTWeakarrow {\tau} 100}
        $$\forall n\ge 0$

@@

$
\begin{array}{ll}
Post_{6a}= post_{4}\shortotimes\ post_{456*}\shortotimes\ post_{6} =&  (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}, \texttt{s2\_ec}\gets \texttt{s1\_ec})\,
\shortotimes \\ &(\texttt{s1\_ec}\gets \texttt{s1\_ec}+n) 
\shortotimes (\texttt{s1\_ec}\gets \texttt{s2\_ec}+1) =\\
&(\texttt{m1\_msg}\!\gets \texttt{s1\_msg}, \texttt{m1\_ec}\!\gets \texttt{s2\_ec}+1\!+\!n, \texttt{s2\_ec}\!\gets \texttt{s2\_ec}+1\!+\!n)
\end{array}$\\

$ WI_{6a}(n) = \openrule
         {\{\}, True, 
         (\texttt{m1\_ec}\gets \texttt{s1\_ec}+1+n, \texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 210} \forall n\ge 0
         $

\medskip
$Post_{8c}= post_{8}\shortotimes\ post_{7}\shortotimes\ post_{4}\shortotimes\ post_{456*} 
= (\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_ec}\gets \texttt{s2\_msg}+1+n), \forall n\ge 0$

 \medskip  
$ WI_{8c}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(s1\_msg,s2\_ec}+n)\}, True, ()}
         {210 \OTWeakarrow {\nounderline{\texttt{out(r1\_msg,r1\_ec)}\subst{(\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_ec}\gets \texttt{s2\_msg}+n)}}} 202}\forall n\ge 0
        $

@@

  $ SS_3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m1)}\}, True,
    (\texttt{b1\_msg}\gets \texttt{m1}, \texttt{b1\_ec}\gets 0)}
  {\text{b0} \OTarrow {\nounderline{\texttt{in(m1)}}} \text{b1}}
  $

>>>>

\subsubsection*{Substitutions}
\label{def:substitutions}

@@

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, V, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of states.
\item[$\bullet$]
$s_0 \in S$ is the initial state.
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
\item[$\bullet$] $V$ is a set of global variables for the pLTS,
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
where $\alpha \in\AlgA$ is a parameterised action, $e_b \in
\AlgB$ is a guard, and the variables $x_j$ 
are assigned the expressions $e_j\in \AlgE$.
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then 
% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! V$, 
		$\vars(e_b)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
		$\forall j\!\in\! J .\,\left(\vars(e_j)\!\subseteq\! V\cup\iv(\alpha)\land 
		x_j\!\in V \right)$. %,  and $s= s'\Rightarrow J=\emptyset$.

@@

We now define the structure of pNets, the following definition relies on the definition 
of holes, leaves and sorts formalised below in Definition~\ref{def-sortholeleave}. Informally, holes are process parameters, leaves provide the set of pLTSs at the leaves of the hierarchical structure of a pNet, and sorts give the signature of a pNet, i.e. the actions it exposes.

@@

Fig. \ref{SimpleProt:Impl} shows the pNet model of the protocol implementation. When the $\symb{Medium}$ detects an error (modelled by a local $\tau$ action), it sends back a $\texttt{m\_error}$ message to the $\symb{Sender}$. The $\symb{Sender}$ increments its local error counter $ec$, and resends the message (including $ec$) to the $\symb{Medium}$, that will, eventually, transmit $m,ec$ to the $\symb{Receiver}$. 
%Strictly speaking, as variables are local to each  state, we have two distinct $ec$ variables in the \textit{Sender} pLTS, one in state \texttt{s1}, another one in state \texttt{s2}, the value of the variable is updated along the transitions entering each of the two states.

@@

\subsection{Open Automata}
 Open automata (OA) are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on a set of variables (potentially with only symbolic values). 
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} (OT) over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$, a set $V$ of variables, and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq J$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action of this open transition.
%        We call \emph{source variables} of an OT the union of all variables
%        in the different terms
%        $\beta_j$ and $\alpha$, and the state variables of $s$.
        \Pred\ is a predicate over the  variables in $V$ and all variables
        in the different terms
       $\beta_j$ and $\alpha$. \Post\ is a set of 
	assignments that are effective \emph{after the open transition}, they are
        represented as a substitution of the form $({x_k\gets e_k})^{k\in K}$ 
	where $\forall k.\, x_k\in V$, and $e_k$ are expressions
        over the variables $V$ and all variables
        in the different terms
       $\beta_j$ and $\alpha$. Because the variables in $V$ can be in both side of the assignments, the assignments are always applied in a parallel manner. Open transitions are identified
        modulo logical equivalence on their predicate. 
\end{definition}

@@

We take in this article a semantics and logical understanding of these
automata. Open automata are closed by a simple form of refinement that
allows us to refine the predicate, or substitute any free variable by
an expression. More formally, let $\Pred$ be any predicate and $\Post$ any substitution such that $V\cap \dom(\Post)=\emptyset$. Then we have the following
implication:

@@

%\LUDO{I added this paragraph}
Because of the semantic interpretation of open automata, the set of open transition of an open automaton is infinite (for example because every free variable can be renamed). However an open automaton is characterized by a  subset of these open transition which is sufficient to generate, by substitution the other ones. In the following, we will abusively write that we define an ``open automaton'' when we provide only the set of open transitions that is sufficient to generate a proper open automaton by saturating each open transition by all possible substitutions and refinements.

@@

We first define states of open pNets as tuples of states. We denote them
 as $\triangleleft\ldots\triangleright$ for distinguishing tuple 
states from other tuples.
\begin{definition}[States of open pNets]\label{def-states}
  A state of an open pNet is a tuple (not necessarily finite) of the
  states of its leaves.

@@

Somehow, this predicate entails a verification of satisfiability in the sense that if the 
predicate $\Predsv$ is not satisfiable, then the transition associated with the 
synchronisation will not occur in the considered state, or will occur with a \False\ precondition which is equivalent.
If the action families do not match or if there is no valuation of
variables such that the above formula can be ensured then the predicate is undefined.

@@

 \smallskip\noindent
 \[  OT_1  = \openrule{
      \{\texttt{P}\mapsto \texttt{p-send(m)}\},  [\texttt{m=m'}],
        \{\texttt{s\_msg}\gets \texttt{m}\}
                      }
    {\ostate{s_0,m_0,r_0} \OTarrow{\nounderline{\texttt{in(m')}}} \ostate{s_1,m_0,r_0}}
    \]

    \smallskip
    The global states here are triples, build as the product of states of the 3 pLTS (remember the holes have no state). The assignment
    $Post$ uses the variable $\texttt{m}$ from the action of hole $\texttt{P}$ to set the
    value of the sender variable named $\texttt{s\_msg}$.

@@

 %% \smallskip\noindent
 %% $  OT_2  = \openrule{
 %%   \emptyset; Pred=[send(SV\_m)=send(s_\_m);in(SV\_m)=in(?m)];  \\ Post=\{m_\_m:=?m\}
 %%                      }
 %%    {\ostate{s_1,m_0,r_0} \OTarrow{\nounderline{send(SV\_m)}} \ostate{s_2,m_1,r_0}}
 %%    $

@@

 %%     $  OT_2  = \openrule{
 %%      \emptyset; Pred=\True;  \\ Post=\{m_\_m:=s_\_m\}
 %%                      }
 %%    {\ostate{s_1,m_0,r_0} \OTarrow{\nounderline{send(s_\_m)}} \ostate{s_2,m_1,r_0}}
 %%    $

 %%    \smallskip
 %%    Note that the global states shows how the sender and the
 %%    first medium have evolved during this transition, the sender
 %%    moving from $s_1$ to $s_2$, and the first (forward) medium from
 %%    $m_0$ to $m_1$.
 %%    The predicate $Pred$ involves the state variables $s_\_m$ and
 %%    $s_\_sb$ of the sender, while the assignment $Post$ sets the
 %%    state variables of the medium.

@@

@@

\subsection*{Example}  Figure~\ref{SimpleProtCounter:SpecOA} shows the open automaton computed from the Specification pNet given in Figure \ref{SimpleProt:Spec}. 
For later references, we give to the transitions of this (strong)
Specification automaton as $SS_i$ while transitions of the
Implementation automaton will be labelled $SI_i$. In the figure we
have annotated each global state with the set of corresponding pLTS variables.

@@

\end{definition}
Rule $\WTUn$ states that it is always possible to do a null $\tau$ transition, where the state is unchanged and the holes perform no action. Rule~$\WTDeux$ states that each open transition can be considered as a weak open transition. The last rule is the most interesting:  Rule~$\WTTrois$ allows any number of $\tau$ transitions before or after any weak open transition. This rules carefully composes predicates, effects, and actions of the holes, indeed in the rule, predicate $\Pred_2$ manipulates variables of $s_1$ that result from the first weak open transition. Their values thus depend on the initial state but also on the effect (as a substitution $\Post_1$) of the first weak open transition. In the same manner, $\Pred_3$ must be applied the joint substitution $\Post_2\bigotimes\Post_1$. Similarly, effects on variables must be applied to obtain the global effect of the composed weak open transition, it must also be applied to observable actions of the holes, and to the global action of the weak open transition.

@@

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.  For readability, we abbreviate the weak open transitions encoded by $\openrule   {\{\}, True,	() } {s \OTWeakarrow \tau s'}$  as $W_\tau$. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds a $WT_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+1) }
         {\text{b1}\OTWeakarrow \tau \text{b1}}$; $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+1) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$;  $\openrule
         {\{\}, True,
			() }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{b\_ec}\gets \texttt{b\_ec}+2) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
        {\text{b1} \OTWeakarrow \tau \text{b1}} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{\nounderline{out(b\_m,b\_ec)}}  as $\alpha$, we get the resulting weak open transition between b1 and b0 (as shown in Figure \ref{WOT2}).  Applying the substitutions finaly yields the weak transition $WS_7$ in Figure  \ref{SimpleProtCounter:WeakSpecOA}.

@@

 For readability, we do not include the full details of the Implementation weak open transitions in Figure \ref{SimpleProtCounter:ImplWOA2}, but we show some examples below and the full list is included in Appendix \ref{}.
First, let us point out that the weak OT loops ($WI_1$,$WI_2$ and $W_\tau$) on state ${000}$ are also present in all other states, we did not repeat them. Then many WOTs are similar, and numbered accordingly as 3, 3a, 3b, 3c and 8, 8a, 8b, 8c respectively: they only differ by the name of the state variables in their respective source or target states; the "variant" WOTs appear in blue in   Figure \ref{SimpleProtCounter:ImplWOA2}.
\end{example}

@@

Now let us give some details about the construction of the weak automaton of the implementation pNet, obtained by application of the weak rules as explained above. We concentrate on weak open transitions $WI_3$ and $WI_4$. Let us denote as $post_n$ the effect (= substitution) of the strong open transitions $SI_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

$post_3 = (\texttt{s\_msg}\gets \texttt{m}, \texttt{s\_ec}\gets 0)$

$post_4 = (\texttt{m\_msg}\gets \texttt{s\_msg}, \texttt{m\_ec}\gets \texttt{s\_ec})$

@@

$post_6 = (\texttt{s\_ec}\gets \texttt{s\_ec+1})$

@@

$post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (\texttt{s\_ec}\gets \texttt{s\_ec}+1)$.
\footnote{when showing  the result of $Posts$ composition, we will systematicaly omit the identity substitutions introduced bythe $\shortotimes$ definition in page \pageref{def:substitutions}}
\medskip

So if we denote ${post_{456*}}$ any iteration of this loop, we get ${post_{456*}} = (\texttt{s\_ec}\gets \texttt{s\_ec}+n)$ for any $n\ge 0$, and the \Post~ of the weak OT  $WI_{3}$ is:\\
 $Post_3 = post_{456*}\shortotimes\ post_3 = (\texttt{s\_msg}\gets m, \texttt{s\_ec}\gets n), \forall n\ge 0$ and \Post~ of  $WI_{3a}$  is:\\ $post_4\shortotimes {post_{456*}}\shortotimes\ post_3 = (\texttt{m\_msg}\gets m, \texttt{m\_ec}\gets n), \forall n\ge 0$.
\medskip

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s\_msg}\gets \texttt{m}, \texttt{s\_ec}\gets n)}
  {000 \OTWeakarrow {\nounderline{\texttt{in(m)}}} 100}
        $$\forall n\ge 0$

$ WI_4(n) = \openrule
         {\{\}, True, 
   (\texttt{m\_msg}\!\gets\! \texttt{s\_msg}, \texttt{m\_ec}\!\gets\! \texttt{s\_ec}+n, \texttt{s\_ec}\!\gets\! \texttt{s\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
        $$\forall n\ge 0$

$ WI_{456}(n) = \openrule
         {\{\}, True, 
    (\texttt{s\_ec}\gets \texttt{s\_ec}+n)}
  {100 \OTWeakarrow {\tau} 100}
        $$\forall n\ge 0$

@@

$
\begin{array}{ll}
Post_{6a}= post_{4}\shortotimes\ post_{456*}\shortotimes\ post_{6} =&  (\texttt{m\_msg}\gets \texttt{s\_msg}, \texttt{m\_ec}\gets \texttt{s\_ec})\,
\shortotimes \\ &(\texttt{s\_ec}\gets \texttt{s\_ec}+n) 
\shortotimes (\texttt{s\_ec}\gets \texttt{s\_ec}+1) =\\
&(\texttt{m\_msg}\!\gets \texttt{s\_msg}, \texttt{m\_ec}\!\gets \texttt{s\_ec}+1\!+\!n, \texttt{s\_ec}\!\gets \texttt{s\_ec}+1\!+\!n)
\end{array}$\\

$ WI_{6a}(n) = \openrule
         {\{\}, True, 
         (\texttt{m\_ec}\gets \texttt{s\_ec}+1+n, \texttt{s\_ec}\gets \texttt{s\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 210} \forall n\ge 0
         $

\medskip
$Post_{8c}= post_{8}\shortotimes\ post_{7}\shortotimes\ post_{4}\shortotimes\ post_{456*} 
= (\texttt{r\_msg}\gets \texttt{s\_msg}, \texttt{r\_ec}\gets \texttt{s\_msg}+1+n), \forall n\ge 0$

 \medskip  
$ WI_{8c}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(s\_msg,s\_ec}+n)\}, True, ()}
         {210 \OTWeakarrow {\nounderline{\texttt{out(r\_msg,r\_ec)}\subst{(\texttt{r\_msg}\gets \texttt{s\_msg}, \texttt{r\_ec}\gets \texttt{s\_msg}+n)}}} 202}\forall n\ge 0
        $

@@

  $ SS_3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m1)}\}, True,
    (\texttt{b\_msg}\gets \texttt{m1}, \texttt{b\_ec}\gets 0)}
  {\text{b0} \OTarrow {\nounderline{\texttt{in(m1)}}} \text{b1}}
  $
